---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Code used for the talk

Packages needed

```{r, message=FALSE, eval=TRUE}
library(tidyverse)
library(paletteer)
library(farver)
library(colorspace)
library(colorscience)
library(recipes)
library(pdist)
```

We source in 3 files. First file compiles the initial dataset with colors as a list-column. Second file includes the hand-engineered feature functions. Last file is a custom plotting function taken from https://github.com/EmilHvitfeldt/r-color-palettes.

```{r, eval=TRUE}
source("R/data_list.R")
source("R/feature_functions.R")
source("R/list_plotter.R")
```

Looking at the initial data

```{r, eval=TRUE}
head(data_list)
```

Here we combine `mutate` with some *purrr* magic to apply the feature functions to the data.

```{r}
data_num <- data_list %>%
  mutate(# contains all a certain color
         contains_yellow = map_dbl(colors, ~ color_contains(.x, "yellow", "hsv")),
         contains_orange = map_dbl(colors, ~ color_contains(.x, "orange", "hsv")),
         contains_red = map_dbl(colors, ~ color_contains(.x, "red", "hsv")),
         contains_purple = map_dbl(colors, ~ color_contains(.x, "purple", "hsv")),
         contains_blue = map_dbl(colors, ~ color_contains(.x, "blue", "hsv")),
         contains_green = map_dbl(colors, ~ color_contains(.x, "green", "hsv")),
         contains_brown = map_dbl(colors, ~ color_contains(.x, "brown", "hsv")),
         contains_white = map_dbl(colors, ~ color_contains(.x, "white", "hsv")),
         contains_black = map_dbl(colors, ~ color_contains(.x, "black", "hsv")),
         # All a certain color
         all_contains_yellow = map_dbl(colors, ~ color_all_contains(.x, "yellow", "hsv")),
         all_contains_orange = map_dbl(colors, ~ color_all_contains(.x, "orange", "hsv")),
         all_contains_red = map_dbl(colors, ~ color_all_contains(.x, "red", "hsv")),
         all_contains_purple = map_dbl(colors, ~ color_all_contains(.x, "purple", "hsv")),
         all_contains_blue = map_dbl(colors, ~ color_all_contains(.x, "blue", "hsv")),
         all_contains_green = map_dbl(colors, ~ color_all_contains(.x, "green", "hsv")),
         all_contains_brown = map_dbl(colors, ~ color_all_contains(.x, "brown", "hsv")),
         all_contains_white = map_dbl(colors, ~ color_all_contains(.x, "white", "hsv")),
         all_contains_black = map_dbl(colors, ~ color_all_contains(.x, "black", "hsv")),
         # Is it linear in a perceptually uniform space
         linear = map_dbl(colors, ~ linear(.x, "hunterlab")),
         linear_deutan = map_dbl(colors, ~ linear(deutan(.x), "hunterlab")),
         linear_protan = map_dbl(colors, ~ linear(protan(.x), "hunterlab")),
         linear_tritan = map_dbl(colors, ~ linear(tritan(.x), "hunterlab")),
         # Twice linear
         twice_linear = map_dbl(colors, ~ linear_split(.x, "hunterlab")),
         twice_linear_deutan = map_dbl(colors, ~ linear_split(deutan(.x), "hunterlab")),
         twice_linear_protan = map_dbl(colors, ~ linear_split(protan(.x), "hunterlab")),
         twice_linear_tritan = map_dbl(colors, ~ linear_split(tritan(.x), "hunterlab")),
         # Min distance between points
         min_distance = map_dbl(colors, ~ min_distance(.x, "hunterlab")),
         min_distance_deutan = map_dbl(colors, ~ min_distance(deutan(.x), "hunterlab")),
         min_distance_protan = map_dbl(colors, ~ min_distance(protan(.x), "hunterlab")),
         min_distance_tritan = map_dbl(colors, ~ min_distance(tritan(.x), "hunterlab")),
         # Max distance between points
         max_distance = map_dbl(colors, ~ max_distance(.x, "hunterlab")),
         max_distance_deutan = map_dbl(colors, ~ max_distance(deutan(.x), "hunterlab")),
         max_distance_protan = map_dbl(colors, ~ max_distance(protan(.x), "hunterlab")),
         max_distance_tritan = map_dbl(colors, ~ max_distance(tritan(.x), "hunterlab")),
         # IQR distance between points
         iqr_distance = map_dbl(colors, ~ iqr_distance(.x, "hunterlab")),
         iqr_distance_deutan = map_dbl(colors, ~ iqr_distance(deutan(.x), "hunterlab")),
         iqr_distance_protan = map_dbl(colors, ~ iqr_distance(protan(.x), "hunterlab")),
         iqr_distance_tritan = map_dbl(colors, ~ iqr_distance(tritan(.x), "hunterlab"))
         )
```

Creating dummy variables and scaling.

```{r}
data_meta <- data_num %>%
  mutate(id = as.character(str_glue("{package}-{palette}-{group}")))

data_prescaled <- data_meta %>%
  select(-package, -palette, -colors) %>%
  mutate_at(vars(type, group), as.factor)

re <- recipe(id ~ ., data = data_prescaled) %>%
  step_dummy(type, group) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  prep()

data_scaled <- bake(re, new_data = data_prescaled)
```

saving data.

```{r}
fs::dir_create("data")
write_csv(data_scaled, "data/data_scaled.csv")
write_rds(data_meta, "data/data_meta.rds")
```

Reading data.

```{r, message=FALSE, eval=TRUE}
data_scaled <- read_csv("data/data_scaled.csv")
data_meta <- read_rds("data/data_meta.rds")
```

Plotting function, takes the arguement pick and n. Pick must be "guess", "random" or "user". See examples below for use.

```{r, eval=TRUE}
example <- function(pick = "guess", n = 9, pal) {
  
  sample <- data_scaled %>% 
    sample_n(1) %>%
    mutate(id = as.character(id))
  
  if (pick == "user") {
    sample <- tibble(id = paste(pal$package, 
                                pal$palette, 
                                pal$group, 
                                sep = "-", collapse = "")) %>%
      left_join(data_scaled, by = "id")
  }
  
  if (pick %in% c("guess", "user")) {
    guesses <- pdist(X = select(data_scaled, -id), Y = select(sample, -id))@dist %>%
      tibble(sum = .,
             id = data_meta$id) %>% 
      arrange(sum) %>% 
      slice(seq_len(n + 1))
  }
  if (pick == "random") {
    guesses <- tibble(id = sample(data_meta$id, n))
  }
  
  main <- left_join(sample, data_meta, by = "id") %>%
    select(colors, package, palette, group)
  
  full_guesses <- left_join(guesses, data_meta, by = "id") %>% slice(-1)
  
  print(main %>%
    mutate(out = str_glue("paletteer_{group}(\"{package}::{palette}\")")) %>%
    pull(out))
  print("------------")
  print(full_guesses %>%
    mutate(out = str_glue("paletteer_{group}(\"{package}::{palette}\")")) %>%
    pull(out))
  
  plot_colors <- c(main %>% pull(colors), 
                   c("#FFFFFF"), 
                   left_join(guesses, data_meta, by = "id") %>% 
                     slice(-1) %>% 
                     pull(colors))
  
  list_plotter(plot_colors, names = "", package_name = "")
}


```

# Examples

```{r, eval=TRUE, fig.height=5, fig.width=4}
example()
```

```{r, eval=TRUE, fig.height=5, fig.width=4}
example(pick = "random")

```

```{r, eval=TRUE, fig.height=5, fig.width=4}
my_pick <- data_list %>%
  filter(package == "wesanderson", palette == "Moonrise2")
example(pick = "user", 9, my_pick)
```

